Market-based Mechanisms for Resource Scheduling

### 2013-12-18 paper/paper.tex
format edits

\subsection{Bidding policy}
Within our setting, the framework must associate a willingness to pay with each
task it wishes to run.

\paragraph{Default policy} For backwards compatibility, 
and as a starting point for designing more complex policies, we want to be able to
use existing Mesos frameworks which are not aware of prices. For these, we
define a default bidding policy which assumes that all tasks are equally
important.

@@

Formally, each framework calculates the current net present value
of their accumulated wealth together with their current obligations
and stream of income, and then bids a price proportional
to this total. This guarantees that the frameworks will
maintain roughly equal budgets (and hence use roughly equal shares
of the cluster), since any framework which became richer
than the others would begin winning auctions.

\paragraph{Adjusting bids} To modify the 
default policy to accomodate inhomogeneous resources,
times, or tasks, we can simply modify framework's
bids to adjust for the desirability of the resource being
purchased or the importance of the purpose they are being used for.

For example, if a task will take less time when run on a machine
which already has relevant data loaded into memory,
a framework can increase its willingness to pay for resources on that machine
by an amount depending on the anticipated speedup.
%XXX mention why we don't care about any properties other than those that we mention

\paragraph{Defensive bidding} One negative consequence of using a closed economy,
in which the frameworks bid using currency which is entirely internal to the cluster,
%XXX mention this distinction elsewhere, and the fact that we can do either
is the possibility of \emph{hoarding}, in which one framework
declines to use any resources during a certain period so that it can amass
a much larger budget than any other resources.
It can then spend this budget at a future point in order to monopolize the cluster,
which may have undesirable conseuqences for frameworks that need uninterrupted service.

@@

We say that a system is ``demand-based'' if it elicits a demand from each user and makes assignments on the basis of the stated demands. For example, DRF, asset allocation, and CEEI all fall into this class.

\begin{theorem}\label{demand-bad}
Any demand-based system which is Pareto efficient under realistic conditions must sometimes assign the entire cluster to a single user, an extreme violation of fairness.
\end{theorem}

\begin{pproof}
We provide the proof for the time-varying demands case. The other arguments are exactly analogous. We give an extremely simple and artificial case, though analogous behavior would be expected in more realistic cases.

@@

But now consider the alternative allocation which exchanges these periods, i.e. assigns each user R resources more during their high priority period and R resources fewer during their low priority period. By definition, this alternative allocation is preferred to the real allocation, contradicting Pareto efficiency.
\qed
\end{pproof}

\begin{theorem}
Under realistic conditions, the implementation of DRF in Mesos is not strategy-proof, Pareto efficient, envy-free, or compatible with the sharing incentive. 
\end{theorem}

DRF satisfies a weaker form of strategy-proofness which we call ``local strategy-proofness,'' 
such that a user is incentivized to honestly state their demands given their beliefs about the effect of those demands on the scheduler's future behavior. 
This differs from the ordinary notion in that a user may need to reason strategically about future rounds of allocation in order to determine
their demand, rather than honestly reporting their real demand.
%XXX replace user by framework
%XXX fold this into the clearer general discussion of loca guarantees

Proof: For variety, we consider the case of inhomogeneous resources. Again, the proof is identical in any of the given cases. For each property we construct a special situation in which DRF fails to achieve that property.

\begin{itemize}
\item \textbf{Strategy-proofness:} consider the setting where a framework's top priority is achieving some particularly valuable resource X 
which is currently in use, though they have other tasks which they value less. To be most extreme, suppose the user values any outcome in which they receive X 
higher than any outcome in which they do not. 
A user can arrange to receive X by strategically passing up resource offers which they could use, in order to maximize 
their probability of being offered the resource X when it first becomes free. 
\item \textbf{Pareto-efficiency:} This is a corollary of theorem~\ref{demand-bad}. 
In the case of inhomogeneous resources, suppose that two users have streams of tasks, some of which run faster on resources of type A and some of which run faster on resources of type B. 
Then no demand-based scheduler can ensure that tasks are always placed on the resources on which they run fastest.
\item \textbf{Envy-freeness and compatibility with sharing incentive:} Suppose that one user consumes $\frac 13$ of the resources in the cluster and another consumes $\frac 12$ of the resources, 
but a particular $\frac 1{10}$ of the resources in the cluster are $10$ times more valuable (to every user) than they appear. 
This could be based on hardware characteristics to which the scheduler is oblivious. 
The user who consumes $\frac 13$ of the resources receives all resource offers first, and so claims all of the especially valuable resources. 
But this means that this user receives significantly more total value than the other user, violating both envy-freeness and the sharing incentive. 
\end{itemize}

The results above hold only for demand-based policies.
The following result shows that the strong guarantees of the sharing
incentive and Pareto efficiency are actually incompatible in general.
This motivates us to introduce weaker notions in the next section.
\begin{theorem}\label{no-efficiency}
%XXX change the appearance of "theorem 3" below to reference this one
The sharing incentive and Pareto efficiency are incompatible under realistic conditions.
\end{theorem}
\begin{pproof}
Consider a setting in which there are two phases and two users. 
The second phase differs from the first in one of two ways: 
either it is the case that user 2 cares more about their tasks in phase 2 than about their tasks in phase 1, 
or it is the case that user 1 cares more about their tasks in phase 2 than about their tasks in phase 1. 
By ``cares more about'' we mean simply that a user would, all else equal, 
prefer to complete those tasks more rapidly at the expense of completing other tasks less rapidly.

If user $i$ values jobs more in phase 2, it is easy to verify that any allocation which is both Pareto efficient and compatible with the sharing incentive must give user $i$ 
more resources than user $j$ in phase 2 and (consequently) fewer resources than user $j$ in phase 1. 
Since the scheduler must decide which user to give more resources in phase 1, it cannot possibly be Pareto efficient and compatible with the sharing incentive in both cases.
\end{pproof}

@@

We define a user's fair share of the cluster to be a proportional share of each resource, with the proportion determined by the ratio between the total amount paid by that user, and the amount paid by all users over the cluster.

\begin{theorem}[Sharing incentive]
Over the life of the cluster, each user A values the total resources they receive at least as much as they value their fair share of the cluster. 
\end{theorem}

Note that users may value resources at different times differently, and this is a statement across the entire lifetime of the cluster.
\begin{pproof}
Consider a particular user A. 
During any auction, let $p_A$ be the price that A pays, let $p_{-A}$ be the price paid by all other users, let $R_A$ be the resources received by A, 
and let $R_{-A}$ be the resources received by all other users. The key point is that user A values $R_A$ at at least $p_A$, and values $R_{-A}$ at at most $p_{-A}$.

Summing over all rounds of the auction, A values the total resources they receive at at least the amount they pay for them, 
and user A values the total resources others receive at at most the amount that others pay for them. The desired result follows immediately.
\end{pproof}

@@

%XXX link this into the general discussion
As with the envy-free analysis of DRF, we need to assume that either there are no complementarities between resources, or that the user's demand for a resource takes into account their expectations about future allocations (e.g. if resources R and Q are only valuable together, we either assume that the scheduler packages them together into a single resource or that the frameworks are able to evaluate the value of r based on the probability that they will be able to receive Q).

\begin{theorem}[Local envy-freeness]
No user who bids honestly would prefer the allocation of another user who paid no more than they did.
\end{theorem}

\begin{pproof}
We will show by induction that for any number of rounds, user A would prefer to have received user A's allocation at user A's price than user B's allocation at user B's price. Consider a particular round, after which user A receives some resources $R_A$ at a price $p_A$ and user B receives resources $R_B$ at a price $p_B$.

By the definition of our auction, the price $p_B$ exceeds user A's willingness to pay for resources $R_B$ 
(or else user A's proxy would have continued bidding and the auction would not have concluded), 
while the price $p_A$ is less than user A's willingness to pay for resources $R_A$ 
(or else user A's proxy would have stopped bidding earlier). 
In particular, user A prefers their package to user B's package. 
Since at the start of the round user A preferred their allocation at their price to user B's allocation at user B's price, 
this property also holds at the end of the round. 
In this step we rely on our assumption that user A's valuations for their own holdings do not change after the auction is announced 
(otherwise it may be possible for user A to become envious in between rounds as more information is revealed, rather than during a round of the auction). 
We note in passing that similar assumptions are required to prove that DRF is envy-free in the dynamic setting,
even in a simple rather than realistic enviornment.
\qed
\end{pproof}

%XXX insert discussion of strategy-proofness of DRF, which mentions (1) the fact that you need to ignore complexities of items / trust users to incorporate them into their bids, (2) the fact that you need to ignore the consequences of your behavior for other participants

\begin{theorem}[Local strategy-proofness] 
Assuming that users are indifferent to changes in each others' budgets,
and that users' willingness to pay 
incorporates accurate expectations of effects on future allocations, 
the optimal behavior in our auction is to honestly declare willingness to pay.
\end{theorem}

\begin{pproof}
We give the standard argument for the dominant strategy truthfulness (strategy-proofness) of Vickrey auctions. 
Indeed, suppose on the contrary 
that $$B_1$$ and $B_2$ are two different bids for user A; that $R_1$ and $R_2$ are the resources received after making either of these bids and $p_1$, $p_2$ are the associated prices; that $B_1$ is an honest bid; and that user A prefers $(R_2, p_2)$ to $(R_1, p_1)$.

@@

This contradicts the assumption that $(R_1, p_1)$ was the result of running the auction with bid $B_1$. 
\end{pproof}

We must also weaken the notion of Pareto efficiency in order for it to be achievable (see Theorem 3)

%XXX make notes about approximations, say we are analyzing the intractable optimization and that statements hold with appropriate caveats, e.g. you aren't Pareto efficient if you can't do the optimization exactly, but you are as Pareto efficient as you can realistically be. Since optimization is computationally hard, the only way we can avoid this caveat is by ignoring it.]
%XXX make notes about the notion of ``local'' guarantees

\begin{theorem}[Local Pareto efficiency.]
If users bid honestly, 
there is no way to change the tasks which are scheduled at a single time $T$ in a way which makes every user better off.
\end{theorem}

\begin{pproof}
At each individual time $T$, the allocation selected is chosen precisely to maximize welfare using the stated preferences of bidders. 
Consequently, assuming that bidders honestly report their welfare, the allocation which is selected maximizes welfare.
\end{pproof}

@@

To evaluate the performance of the market-based mechanism compared with DRF,
we implemented an event-based simulation of Mesos with both resource scheduling algorithms.
We study the performance of DRF and the market-based mechanism both when pricing information is not available and a default bidding policy must be used, and when pricing information is specified under a specific policy. These simulations were run under a statistical workload generated from a month-long Google cluster trace \cite{TODO}.

\subsection{Simulation Implementation\\}

Our evaluation is conducted through an event-based simulator of Mesos, implemented with both the DRF and the auction-based resource scheduling algorithms. We chose to use simulations because it allowed for simplified evaluation over a variety of environmental variables, such as the number of nodes in the cluster or resources per node, and provided an efficient method for testing various bidding policies. Our simulator was implemented in C++ using approximately 1200 lines of code.

@@

\subsection{Performance Without Pricing\\}

@@

\subsection{Performance With Pricing\\}

When pricing information is provided, we show that the market-based mechanism can provide better performance than in DRF. As discussed in Section TODO, there are many situations in which prices can be used to better encapsulate a framework's preferences. As a case study, we investigate one particular scenario of inhomogeneous performing resources.

@@

>>>>

\subsection{Bidding policy}
Within our setting, the framework must associate a willingness to pay with each
task it wishes to run. 
\vspace{-5mm}
\paragraph{Default policy} For backwards compatibility, 
and as a starting point for designing more complex policies, we want to be able to
use existing Mesos frameworks which are not aware of prices. For these, we
define a default bidding policy which assumes that all tasks are equally
important.

@@

Formally, each framework calculates the current net present value
of their accumulated wealth together with their current obligations
and stream of income, and then bids a price proportional
to this total. This guarantees that the frameworks will
maintain roughly equal budgets (and hence use roughly equal shares
of the cluster), since any framework which became richer
than the others would begin winning auctions.
\vspace{-5mm}
\paragraph{Adjusting bids} To modify the 
default policy to accomodate inhomogeneous resources,
times, or tasks, we can simply modify framework's
bids to adjust for the desirability of the resource being
purchased or the importance of the purpose they are being used for.

For example, if a task will take less time when run on a machine
which already has relevant data loaded into memory,
a framework can increase its willingness to pay for resources on that machine
by an amount depending on the anticipated speedup.
%XXX mention why we don't care about any properties other than those that we mention
\vspace{-5mm}
\paragraph{Defensive bidding} One negative consequence of using a closed economy,
in which the frameworks bid using currency which is entirely internal to the cluster,
%XXX mention this distinction elsewhere, and the fact that we can do either
is the possibility of \emph{hoarding}, in which one framework
declines to use any resources during a certain period so that it can amass
a much larger budget than any other resources.
It can then spend this budget at a future point in order to monopolize the cluster,
which may have undesirable conseuqences for frameworks that need uninterrupted service.

@@

We say that a system is ``demand-based'' if it elicits a demand from each user and makes assignments on the basis of the stated demands. For example, DRF, asset allocation, and CEEI all fall into this class.
\vspace{-5mm}
\begin{theorem}\label{demand-bad}
Any demand-based system which is Pareto efficient under realistic conditions must sometimes assign the entire cluster to a single user, an extreme violation of fairness.
\end{theorem}
\vspace{-5mm}
\begin{pproof}
We provide the proof for the time-varying demands case. The other arguments are exactly analogous. We give an extremely simple and artificial case, though analogous behavior would be expected in more realistic cases.

@@

But now consider the alternative allocation which exchanges these periods, i.e. assigns each user R resources more during their high priority period and R resources fewer during their low priority period. By definition, this alternative allocation is preferred to the real allocation, contradicting Pareto efficiency.
\qed
\end{pproof}
\vspace{-5mm}
\begin{theorem}
Under realistic conditions, the implementation of DRF in Mesos is not strategy-proof, Pareto efficient, envy-free, or compatible with the sharing incentive. 
\end{theorem}
\vspace{-5mm}
DRF satisfies a weaker form of strategy-proofness which we call ``local strategy-proofness,'' 
such that a user is incentivized to honestly state their demands given their beliefs about the effect of those demands on the scheduler's future behavior. 
This differs from the ordinary notion in that a user may need to reason strategically about future rounds of allocation in order to determine
their demand, rather than honestly reporting their real demand.
%XXX replace user by framework
%XXX fold this into the clearer general discussion of loca guarantees

\textbf{Proof} For variety, we consider the case of inhomogeneous resources. Again, the proof is identical in any of the given cases. For each property we construct a special situation in which DRF fails to achieve that property.
\vspace{-5mm}
\begin{itemize}
\itemsep0em
\item \textbf{Strategy-proofness:} consider the setting where a framework's top priority is achieving some particularly valuable resource X 
which is currently in use, though they have other tasks which they value less. To be most extreme, suppose the user values any outcome in which they receive X 
higher than any outcome in which they do not. 
A user can arrange to receive X by strategically passing up resource offers which they could use, in order to maximize 
their probability of being offered the resource X when it first becomes free. 
\item \textbf{Pareto-efficiency:} This is a corollary of theorem~\ref{demand-bad}. 
In the case of inhomogeneous resources, suppose that two users have streams of tasks, some of which run faster on resources of type A and some of which run faster on resources of type B. 
Then no demand-based scheduler can ensure that tasks are always placed on the resources on which they run fastest.
\item \textbf{Envy-freeness and compatibility with sharing incentive:} Suppose that one user consumes $\frac 13$ of the resources in the cluster and another consumes $\frac 12$ of the resources, 
but a particular $\frac 1{10}$ of the resources in the cluster are $10$ times more valuable (to every user) than they appear. 
This could be based on hardware characteristics to which the scheduler is oblivious. 
The user who consumes $\frac 13$ of the resources receives all resource offers first, and so claims all of the especially valuable resources. 
But this means that this user receives significantly more total value than the other user, violating both envy-freeness and the sharing incentive. 
\end{itemize}

The results above hold only for demand-based policies.
The following result shows that the strong guarantees of the sharing
incentive and Pareto efficiency are actually incompatible in general.
This motivates us to introduce weaker notions in the next section.
\vspace{-5mm}
\begin{theorem}\label{no-efficiency}
%XXX change the appearance of "theorem 3" below to reference this one
The sharing incentive and Pareto efficiency are incompatible under realistic conditions.
\end{theorem}
\vspace{-5mm}
\begin{pproof}
Consider a setting in which there are two phases and two users. 
The second phase differs from the first in one of two ways: 
either it is the case that user 2 cares more about their tasks in phase 2 than about their tasks in phase 1, 
or it is the case that user 1 cares more about their tasks in phase 2 than about their tasks in phase 1. 
By ``cares more about'' we mean simply that a user would, all else equal, 
prefer to complete those tasks more rapidly at the expense of completing other tasks less rapidly.

If user $i$ values jobs more in phase 2, it is easy to verify that any allocation which is both Pareto efficient and compatible with the sharing incentive must give user $i$ 
more resources than user $j$ in phase 2 and (consequently) fewer resources than user $j$ in phase 1. 
Since the scheduler must decide which user to give more resources in phase 1, it cannot possibly be Pareto efficient and compatible with the sharing incentive in both cases. \qed
\end{pproof}

@@

We define a user's fair share of the cluster to be a proportional share of each resource, with the proportion determined by the ratio between the total amount paid by that user, and the amount paid by all users over the cluster. 
\vspace{-5mm}
\begin{theorem}[Sharing incentive]
Over the life of the cluster, each user A values the total resources they receive at least as much as they value their fair share of the cluster. 
\end{theorem}
\vspace{-5mm}
Note that users may value resources at different times differently, and this is a statement across the entire lifetime of the cluster.
\vspace{-5mm}
\begin{pproof}
Consider a particular user A. 
During any auction, let $p_A$ be the price that A pays, let $p_{-A}$ be the price paid by all other users, let $R_A$ be the resources received by A, 
and let $R_{-A}$ be the resources received by all other users. The key point is that user A values $R_A$ at at least $p_A$, and values $R_{-A}$ at at most $p_{-A}$.

Summing over all rounds of the auction, A values the total resources they receive at at least the amount they pay for them, 
and user A values the total resources others receive at at most the amount that others pay for them. The desired result follows immediately. \qed
\end{pproof}

@@

%XXX link this into the general discussion
As with the envy-free analysis of DRF, we need to assume that either there are no complementarities between resources, or that the user's demand for a resource takes into account their expectations about future allocations (e.g. if resources R and Q are only valuable together, we either assume that the scheduler packages them together into a single resource or that the frameworks are able to evaluate the value of r based on the probability that they will be able to receive Q).
\vspace{-5mm}
\begin{theorem}[Local envy-freeness]
No user who bids honestly would prefer the allocation of another user who paid no more than they did.
\end{theorem}
\vspace{-5mm}
\begin{pproof}
We will show by induction that for any number of rounds, user A would prefer to have received user A's allocation at user A's price than user B's allocation at user B's price. Consider a particular round, after which user A receives some resources $R_A$ at a price $p_A$ and user B receives resources $R_B$ at a price $p_B$.

By the definition of our auction, the price $p_B$ exceeds user A's willingness to pay for resources $R_B$ 
(or else user A's proxy would have continued bidding and the auction would not have concluded), 
while the price $p_A$ is less than user A's willingness to pay for resources $R_A$ 
(or else user A's proxy would have stopped bidding earlier). 
In particular, user A prefers their package to user B's package. 
Since at the start of the round user A preferred their allocation at their price to user B's allocation at user B's price, 
this property also holds at the end of the round. 
In this step we rely on our assumption that user A's valuations for their own holdings do not change after the auction is announced 
(otherwise it may be possible for user A to become envious in between rounds as more information is revealed, rather than during a round of the auction). 
We note in passing that similar assumptions are required to prove that DRF is envy-free in the dynamic setting,
even in a simple rather than realistic enviornment.
\qed
\end{pproof}
%XXX insert discussion of strategy-proofness of DRF, which mentions (1) the fact that you need to ignore complexities of items / trust users to incorporate them into their bids, (2) the fact that you need to ignore the consequences of your behavior for other participants
\vspace{-8mm}
\begin{theorem}[Local strategy-proofness] 
Assuming that users are indifferent to changes in each others' budgets,
and that users' willingness to pay 
incorporates accurate expectations of effects on future allocations, 
the optimal behavior in our auction is to honestly declare willingness to pay.
\end{theorem}
\vspace{-7mm}
\begin{pproof}
We give the standard argument for the dominant strategy truthfulness (strategy-proofness) of Vickrey auctions. 
Indeed, suppose on the contrary 
that $B_1$ and $B_2$ are two different bids for user A; that $R_1$ and $R_2$ are the resources received after making either of these bids and $p_1$, $p_2$ are the associated prices; that $B_1$ is an honest bid; and that user A prefers $(R_2, p_2)$ to $(R_1, p_1)$.

@@

This contradicts the assumption that $(R_1, p_1)$ was the result of running the auction with bid $B_1$. \qed 
\end{pproof}
\vspace{-5mm}
We must also weaken the notion of Pareto efficiency in order for it to be achievable (see Theorem 3):
\vspace{-5mm}
%XXX make notes about approximations, say we are analyzing the intractable optimization and that statements hold with appropriate caveats, e.g. you aren't Pareto efficient if you can't do the optimization exactly, but you are as Pareto efficient as you can realistically be. Since optimization is computationally hard, the only way we can avoid this caveat is by ignoring it.]
%XXX make notes about the notion of ``local'' guarantees
\begin{theorem}[Local Pareto efficiency.]
If users bid honestly, 
there is no way to change the tasks which are scheduled at a single time $T$ in a way which makes every user better off.
\end{theorem}
\vspace{-5mm}
\begin{pproof}
At each individual time $T$, the allocation selected is chosen precisely to maximize welfare using the stated preferences of bidders. 
Consequently, assuming that bidders honestly report their welfare, the allocation which is selected maximizes welfare. \qed
\end{pproof}

@@

To evaluate the performance of the market-based mechanism compared with DRF,
we implemented an event-based simulation of Mesos with both resource scheduling algorithms.
We study the performance of DRF and the market-based mechanism both when pricing information is not available and a default bidding policy must be used, and when pricing information is specified under a specific policy. These simulations were run under a statistical workload generated from a month-long Google cluster trace \cite{googlecluster}.
\vspace{-1mm}
\subsection{Simulation Implementation}
Our evaluation is conducted through an event-based simulator of Mesos, implemented with both the DRF and the auction-based resource scheduling algorithms. We chose to use simulations because it allowed for simplified evaluation over a variety of environmental variables, such as the number of nodes in the cluster or resources per node, and provided an efficient method for testing various bidding policies. Our simulator was implemented in C++ using approximately 1200 lines of code.

@@

\subsection{Performance Without Pricing}

@@

\subsection{Performance With Pricing}

When pricing information is provided, we show that the market-based mechanism can provide better performance than in DRF. As a case study, we investigate one particular scenario of inhomogeneous performing resources.

@@

There are a number of related works in the same or similar domains.
